---
title: "BIOSTAT 685 HW1"
author: "Shuo Qin 60109782"
date: "2024-01-28"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(readxl)
library(ggplot2)
df <- read_xlsx(path = "/Users/shuoqin/Desktop/TbHepBLifeExp.xlsx")
df <-  df[df$Year == 2019,]
df_differece <- data.frame(c())
```

## Question 1
### (a)
```{r}
lowincome_19_female <- df[df$Year == 2019 & df$Income == 'Low income'& df$Gender == "Female", ]
lowincome_19_male <- df[df$Year == 2019 & df$Income == 'Low income'& df$Gender == "Male", ]

par(mfrow = c(1,2))
## Histograms
ggplot(data = lowincome_19_female, aes(x = HepatitisB)) +
  geom_histogram() +  
  ggtitle("Histogram of Hepatitis B Female") 


ggplot(data = lowincome_19_male, aes(x = HepatitisB)) +
  geom_histogram()+  
  ggtitle("Histogram of Hepatitis B Male") 
```


### (b)

```{r}
par(mfrow = c(2,1))
## Boxplots

ggplot(data = lowincome_19_female, aes(x = HepatitisB)) +
  geom_boxplot() +  
  ggtitle("Boxplot of Hepatitis B Female") 

ggplot(data = lowincome_19_male, aes(x = HepatitisB)) +
  geom_boxplot() +  
  ggtitle("Histogram of Hepatitis B Male") 


```


### (c)
```{r}
## mean and median
mean_female <- mean(lowincome_19_female$HepatitisB)
median_female <- median(lowincome_19_female$HepatitisB)
mean_male <- mean(lowincome_19_male$HepatitisB)
median_male <- median(lowincome_19_male$HepatitisB)

mean_female
median_female
mean_male
median_male
```



### (d)
```{r}
## t-test
t_test <- t.test(lowincome_19_female$HepatitisB, lowincome_19_male$HepatitisB, var.equal = TRUE)
print(t_test)
```


Based on the result of t-test, we have a p-value of 0.0038 which is less than 0.05. So we reject the null hypothesis which means there's a significant difference in the distribution of the incidence of Hepatitis B between the two gender groups in 2019.



### (e)

```{r}
wilcox.test(lowincome_19_female$HepatitisB, lowincome_19_male$HepatitisB, 
            correct = FALSE, alternative = "two.sided")
```


Based on the result of Wilcoxon rank sum test, we have a p-value of 0.008651 which is less than 0.05. So we reject the null hypothesis which means there's a significant difference in the distribution of the incidence of Hepatitis B between the two gender groups in 2019.



## Question 2
### (a)
```{r}
lm_19_female <- df[df$Year == 2019 & df$Income == 'Lower middle income'& df$Gender == "Female", ]
lm_19_male <- df[df$Year == 2019 & df$Income == 'Lower middle income'& df$Gender == "Male", ]

par(mfrow = c(2,1))
## Histograms
ggplot(data = lm_19_female, aes(x = HepatitisB)) +
  geom_histogram() +  
  ggtitle("Histogram of Hepatitis B Female") 


ggplot(data = lm_19_male, aes(x = HepatitisB)) +
  geom_histogram()+  
  ggtitle("Histogram of Hepatitis B Male") 
```

### (b)

```{r}
par(mfrow = c(2,1))
## Boxplots

ggplot(data = lm_19_female, aes(x = HepatitisB)) +
  geom_boxplot() +  
  ggtitle("Boxplot of Hepatitis B Female") 

ggplot(data = lm_19_male, aes(x = HepatitisB)) +
  geom_boxplot() +  
  ggtitle("Histogram of Hepatitis B Male") 

```


### (c)
```{r}
## mean and median
lm_mean_female <- mean(lm_19_female$HepatitisB)
lm_median_female <- median(lm_19_female$HepatitisB)
lm_mean_male <- mean(lm_19_male$HepatitisB)
lm_median_male <- median(lm_19_male$HepatitisB)

lm_mean_female
lm_median_female
lm_mean_male
lm_median_male
```



### (d)
```{r}
## t-test
t_test <- t.test(lm_19_female$HepatitisB, lm_19_male$HepatitisB, var.equal = TRUE)
print(t_test)
```


Based on the result of t-test, we have a p-value of 0.01255 which is less than 0.05. So we reject the null hypothesis which means there's a significant difference in the distribution of the incidence of Hepatitis B between the two gender groups in 2019.




### (e)

```{r}
wilcox.test(lm_19_female$HepatitisB, lm_19_male$HepatitisB, 
            correct = FALSE, alternative = "two.sided")
```


Based on the result of Wilcoxon rank sum test, we have a p-value of 0.01827 which is less than 0.05. So we reject the null hypothesis which means there's a significant difference in the distribution of the incidence of Hepatitis B between the two gender groups in 2019.


## Question 3
### (a)
```{r}
um_19_female <- df[df$Year == 2019 & df$Income == 'Upper middle income'& df$Gender == "Female", ]
um_19_male <- df[df$Year == 2019 & df$Income == 'Upper middle income'& df$Gender == "Male", ]

par(mfrow = c(2,1))
## Histograms
ggplot(data = um_19_female, aes(x = HepatitisB)) +
  geom_histogram() +  
  ggtitle("Histogram of Hepatitis B Female") 


ggplot(data = um_19_male, aes(x = HepatitisB)) +
  geom_histogram()+  
  ggtitle("Histogram of Hepatitis B Male") 
```


### (b)
```{r}
par(mfrow = c(2,1))
## Boxplots

ggplot(data = um_19_female, aes(x = HepatitisB)) +
  geom_boxplot() +  
  ggtitle("Boxplot of Hepatitis B Female") 

ggplot(data = um_19_male, aes(x = HepatitisB)) +
  geom_boxplot() +  
  ggtitle("Histogram of Hepatitis B Male") 
```


### (c)
```{r}
## mean and median
um_mean_female <- mean(um_19_female$HepatitisB)
um_median_female <- median(um_19_female$HepatitisB)
um_mean_male <- mean(um_19_male$HepatitisB)
um_median_male <- median(um_19_male$HepatitisB)

um_mean_female
um_median_female
um_mean_male
um_median_male
```



### (d)
```{r}
## t-test
t_test <- t.test(um_19_female$HepatitisB, um_19_male$HepatitisB, var.equal = TRUE)
print(t_test)
```


Based on the result of t-test, we have a p-value of 0.03569 which is less than 0.05. So we reject the null hypothesis which means there's a significant difference in the distribution of the incidence of Hepatitis B between the two gender groups in 2019.



### (e)

```{r}
wilcox.test(um_19_female$HepatitisB, um_19_male$HepatitisB, 
            correct = FALSE, alternative = "two.sided")
```


Based on the result of Wilcoxon rank sum test, we have a p-value of 0.008338 which is less than 0.05. So we reject the null hypothesis which means there's a significant difference in the distribution of the incidence of Hepatitis B between the two gender groups in 2019.



### Question 4
### (a)
```{r}
h_19_female <- df[df$Year == 2019 & df$Income == 'High income'& df$Gender == "Female", ]
h_19_male <- df[df$Year == 2019 & df$Income == 'High income'& df$Gender == "Male", ]

par(mfrow = c(2,1))
## Histograms
ggplot(data = h_19_female, aes(x = HepatitisB)) +
  geom_histogram() +  
  ggtitle("Histogram of Hepatitis B Female") 


ggplot(data = h_19_male, aes(x = HepatitisB)) +
  geom_histogram()+  
  ggtitle("Histogram of Hepatitis B Male") 
```


### (b)
```{r}
par(mfrow = c(2,1))
## Boxplots

ggplot(data = h_19_female, aes(x = HepatitisB)) +
  geom_boxplot() +  
  ggtitle("Boxplot of Hepatitis B Female") 

ggplot(data = h_19_male, aes(x = HepatitisB)) +
  geom_boxplot() +  
  ggtitle("Histogram of Hepatitis B Male") 
```


### (c)
```{r}
## mean and median
h_mean_female <- mean(h_19_female$HepatitisB)
h_median_female <- median(h_19_female$HepatitisB)
h_mean_male <- mean(h_19_male$HepatitisB)
h_median_male <- median(h_19_male$HepatitisB)

h_mean_female
h_median_female
h_mean_male
h_median_male
```



### (d)
```{r}
## t-test
t_test <- t.test(h_19_female$HepatitisB, h_19_male$HepatitisB, var.equal = TRUE)
print(t_test)

```


Based on the result of t-test, we have a p-value of 0.01492 which is less than 0.05. So we reject the null hypothesis which means there's a significant difference in the distribution of the incidence of Hepatitis B between the two gender groups in 2019.




### (e)

```{r}
wilcox.test(h_19_female$HepatitisB, h_19_male$HepatitisB, 
            correct = FALSE, alternative = "two.sided")
```


Based on the result of Wilcoxon rank sum test, we have a p-value of 9.802e-05 which is less than 0.05. So we reject the null hypothesis which means there's a significant difference in the distribution of the incidence of Hepatitis B between the two gender groups in 2019.



## Question 5
(a)I would you expect the resulting p-values to be less than the corresponding p-values you got in Questions (1)-(4) for the non-transformed values. Because once we use the log transformation on the right skewed data, we will make it more symmetrical comparing to before and since t-test has the assumption of distribution the p-value will be affected. So it could be less or greater than the result than before. In this dataset and t-test, the resulting p-vlue will be less.


(b)I would you expect the resulting p-values to be equal to the corresponding p-values you got in Questions (1)-(4) for the non-transformed values. Because the wilcoxon sum rank test does not have the assumption about a normal distribution so it will not affect the p-value after the log transformation. 
