---
title: "STATS 506 HW3"
format: html
editor: visual
embed-resources: true
code-fold: true
---

Github Link: 


#### Problem 1(a)

``` stata

 import sasxport5 "K:\STATA 506\VIX_D.XPT", clear
 save "K:\STATA 506\VIX_D.dta"
 import sasxport5 "K:\STATA 506\DEMO_D.XPT", clear
 save "K:\STATA 506\DEMO_D.dta"
 merge 1:1 seqn using "K:\STATA 506\VIX_D.dta", keep(match)
 count
 
```

to produce:

``` stata

. 
end of do-file

. do "C:\Users\LOCAL_~2\Temp\STD2350_000000.tmp"

.  import sasxport5 "K:\STATA 506\DEMO_D.XPT", clear

. 
end of do-file

. do "C:\Users\LOCAL_~2\Temp\STD2350_000000.tmp"

.  save "K:\STATA 506\DEMO_D.dta"
file K:\STATA 506\DEMO_D.dta saved

. 

  do "C:\Users\LOCAL_~2\Temp\STD40c4_000000.tmp"

.  merge 1:1 seqn using "K:\STATA 506\VIX_D.dta", keep(match)

    Result                      Number of obs
    -----------------------------------------
    Not matched                             0
    Matched                             6,980  (_merge==3)
    -----------------------------------------

. count
  6,980


. 
end of do-file

. 
```

## citation:

"https://www.stata.com/manuals/dmerge.pdf"


#### Problem 1(b)

```stata

 recode ridageyr (0/9=0) (10/19=10) (20/29=20) (30/39=30) (40/49=40) (50/59=50) (60/69=60) (70/79=70) (80/89=80) (90/99=90) (100/109=100), generate(ageScale)
 
 tabulate ageScale viq220, row

```

to produce

```stata 

.  recode ridageyr (0/9=0) (10/19=10) (20/29=20) (30/39=30) (40/49=40) (50/59=50
> ) (60/69=60) (70/79=70) (80/89=80) (90/99=90) (100/109=100), generate(ageScale
> )
(1,301 differences between ridageyr and ageScale)

. do "C:\Users\LOCAL_~1\Temp\STD2eec_000000.tmp"

.  tabulate ageScale viq220, row

+----------------+
| Key            |
|----------------|
|   frequency    |
| row percentage |
+----------------+

 RECODE of |
  ridageyr |
   (Age at |
 Screening |
Adjudicate | Glasses/contact lenses worn for
       d - |             distance
   Recode) |         1          2          9 |     Total
-----------+---------------------------------+----------
        10 |       670      1,418          0 |     2,088 
           |     32.09      67.91       0.00 |    100.00 
-----------+---------------------------------+----------
        20 |       306        631          2 |       939 
           |     32.59      67.20       0.21 |    100.00 
-----------+---------------------------------+----------
        30 |       269        481          0 |       750 
           |     35.87      64.13       0.00 |    100.00 
-----------+---------------------------------+----------
        40 |       286        487          0 |       773 
           |     37.00      63.00       0.00 |    100.00 
-----------+---------------------------------+----------
        50 |       335        274          0 |       609 
           |     55.01      44.99       0.00 |    100.00 
-----------+---------------------------------+----------
        60 |       392        238          0 |       630 
           |     62.22      37.78       0.00 |    100.00 
-----------+---------------------------------+----------
        70 |       299        148          0 |       447 
           |     66.89      33.11       0.00 |    100.00 
-----------+---------------------------------+----------
        80 |       208        103          0 |       311 
           |     66.88      33.12       0.00 |    100.00 
-----------+---------------------------------+----------
     Total |     2,765      3,780          2 |     6,547 
           |     42.23      57.74       0.03 |    100.00 

. 
end of do-file

. 
```

## citation :

"https://www.stata.com/manuals/drecode.pdf"
"https://www.stata.com/manuals/rtabulateoneway.pdf"



#### Problem 1(c)
## 1
```stata

 replace viq220 = viq220 - 1
 regress viq220 c.ridageyr
 estat ic
 logit viq220 c.ridageyr, or
 estimates store m1, title(Model 1)

```

to produce

```stata

. do "C:\Users\LOCAL_~1\Temp\STDc3c_000000.tmp"

.  replace viq220 = viq220 - 1
(6,547 real changes made)

. 
end of do-file

. do "C:\Users\LOCAL_~1\Temp\STDc3c_000000.tmp"

.  regress viq220 c.ridageyr

      Source |       SS           df       MS      Number of obs   =     6,547
-------------+----------------------------------   F(1, 6545)      =    446.28
       Model |  108.967394         1  108.967394   Prob > F        =    0.0000
    Residual |  1598.08316     6,545  .244168551   R-squared       =    0.0638
-------------+----------------------------------   Adj R-squared   =    0.0637
       Total |  1707.05056     6,546  .260777659   Root MSE        =    .49413

------------------------------------------------------------------------------
      viq220 | Coefficient  Std. err.      t    P>|t|     [95% conf. interval]
-------------+----------------------------------------------------------------
    ridageyr |   -.005917   .0002801   -21.13   0.000    -.0064661   -.0053679
       _cons |   .8035266   .0122248    65.73   0.000     .7795621    .8274911
------------------------------------------------------------------------------

. 
end of do-file

. do "C:\Users\LOCAL_~1\Temp\STDc3c_000000.tmp"

.  estat ic

Akaike's information criterion and Bayesian information criterion

-----------------------------------------------------------------------------
       Model |          N   ll(null)  ll(model)      df        AIC        BIC
-------------+---------------------------------------------------------------
           . |      6,547  -4889.421  -4673.494       2   9350.988   9364.562
-----------------------------------------------------------------------------
Note: BIC uses N = number of observations. See [R] IC note.

. 
end of do-file

. do "C:\Users\LOCAL_~1\Temp\STDc3c_000000.tmp"

.  logit viq220 c.ridageyr, or

Iteration 0:  Log likelihood = -4458.7243  
Iteration 1:  Log likelihood = -4237.1322  
Iteration 2:  Log likelihood = -4236.8396  
Iteration 3:  Log likelihood = -4236.8396  

Logistic regression                                     Number of obs =  6,547
                                                        LR chi2(1)    = 443.77
                                                        Prob > chi2   = 0.0000
Log likelihood = -4236.8396                             Pseudo R2     = 0.0498

------------------------------------------------------------------------------
      viq220 | Odds ratio   Std. err.      z    P>|z|     [95% conf. interval]
-------------+----------------------------------------------------------------
    ridageyr |    .975618   .0011762   -20.48   0.000     .9733154    .9779259
       _cons |   3.532101   .1887809    23.61   0.000     3.180818     3.92218
------------------------------------------------------------------------------
Note: _cons estimates baseline odds.

. 
end of do-file

. 

```

## 2
```stata

 regress viq220 c.ridageyr i.ridreth1 i.riagendr
 estat ic
 logit viq220 c.ridageyr i.ridreth1 i.riagendr, or
 estimates store m2, title(Model 2)

```

to produce

```stata

 . do "C:\Users\LOCAL_~1\Temp\STDc3c_000000.tmp"

.  regress viq220 c.ridageyr i.ridreth1 i.riagendr

      Source |       SS           df       MS      Number of obs   =     6,547
-------------+----------------------------------   F(6, 6540)      =    106.57
       Model |  152.040298         6  25.3400497   Prob > F        =    0.0000
    Residual |  1555.01026     6,540  .237769153   R-squared       =    0.0891
-------------+----------------------------------   Adj R-squared   =    0.0882
       Total |  1707.05056     6,546  .260777659   Root MSE        =    .48762

------------------------------------------------------------------------------
      viq220 | Coefficient  Std. err.      t    P>|t|     [95% conf. interval]
-------------+----------------------------------------------------------------
    ridageyr |    -.00526   .0002879   -18.27   0.000    -.0058244   -.0046955
             |
    ridreth1 |
          2  |  -.0352318    .037023    -0.95   0.341    -.1078089    .0373453
          3  |  -.1501615   .0158726    -9.46   0.000    -.1812769   -.1190461
          4  |  -.0594469   .0170746    -3.48   0.001    -.0929187   -.0259752
          5  |  -.1487024   .0316435    -4.70   0.000    -.2107341   -.0866707
             |
  2.riagendr |  -.1067196   .0120702    -8.84   0.000    -.1303812    -.083058
       _cons |   .9202886   .0165296    55.68   0.000     .8878851    .9526921
------------------------------------------------------------------------------

.  estat ic

Akaike's information criterion and Bayesian information criterion

-----------------------------------------------------------------------------
       Model |          N   ll(null)  ll(model)      df        AIC        BIC
-------------+---------------------------------------------------------------
           . |      6,547  -4889.421  -4584.053       7   9182.106   9229.614
-----------------------------------------------------------------------------
Note: BIC uses N = number of observations. See [R] IC note.

.  logit viq220 c.ridageyr i.ridreth1 i.riagendr, or

Iteration 0:  Log likelihood = -4458.7243  
Iteration 1:  Log likelihood = -4139.4727  
Iteration 2:  Log likelihood = -4137.9626  
Iteration 3:  Log likelihood = -4137.9624  

Logistic regression                                     Number of obs =  6,547
                                                        LR chi2(6)    = 641.52
                                                        Prob > chi2   = 0.0000
Log likelihood = -4137.9624                             Pseudo R2     = 0.0719

------------------------------------------------------------------------------
      viq220 | Odds ratio   Std. err.      z    P>|z|     [95% conf. interval]
-------------+----------------------------------------------------------------
    ridageyr |   .9776661   .0012342   -17.89   0.000     .9752501     .980088
             |
    ridreth1 |
          2  |    .854417   .1403638    -0.96   0.338     .6192048    1.178977
          3  |   .5122235    .035861    -9.56   0.000     .4465463    .5875603
          4  |   .7688672   .0588752    -3.43   0.001     .6617158    .8933696
          5  |   .5209638   .0705399    -4.82   0.000      .399533    .6793014
             |
  2.riagendr |   .6058933   .0321156    -9.45   0.000     .5461072    .6722246
       _cons |   6.281091   .4894375    23.58   0.000     5.391473      7.3175
------------------------------------------------------------------------------
Note: _cons estimates baseline odds.

. 
end of do-file

. 


```

## 3
```stata

 regress viq220 c.ridageyr i.ridreth1 i.riagendr c.indfmpir
 estat ic
 logit viq220 c.ridageyr i.ridreth1 i.riagendr c.indfmpir, or
 estimates store m3, title(Model 3)

 
 ```
 
 to produce
 
 ```stata
 
.  regress viq220 c.ridageyr i.ridreth1 i.riagendr c.indfmpir

      Source |       SS           df       MS      Number of obs   =     6,249
-------------+----------------------------------   F(7, 6241)      =     88.27
       Model |  147.460482         7  21.0657832   Prob > F        =    0.0000
    Residual |  1489.37997     6,241  .238644444   R-squared       =    0.0901
-------------+----------------------------------   Adj R-squared   =    0.0891
       Total |  1636.84045     6,248  .261978306   Root MSE        =    .48851

------------------------------------------------------------------------------
      viq220 | Coefficient  Std. err.      t    P>|t|     [95% conf. interval]
-------------+----------------------------------------------------------------
    ridageyr |   -.005169   .0002967   -17.42   0.000    -.0057505   -.0045875
             |
    ridreth1 |
          2  |  -.0275455   .0379827    -0.73   0.468    -.1020048    .0469137
          3  |  -.1140486   .0171294    -6.66   0.000    -.1476281   -.0804691
          4  |  -.0488356   .0176895    -2.76   0.006    -.0835132    -.014158
          5  |  -.1234639   .0327039    -3.78   0.000    -.1875748    -.059353
             |
  2.riagendr |  -.1096468   .0123808    -8.86   0.000    -.1339175   -.0853761
    indfmpir |   -.024683   .0040946    -6.03   0.000    -.0327098   -.0166562
       _cons |   .9588157   .0186702    51.36   0.000     .9222156    .9954157
------------------------------------------------------------------------------

.  estat ic

Akaike's information criterion and Bayesian information criterion

-----------------------------------------------------------------------------
       Model |          N   ll(null)  ll(model)      df        AIC        BIC
-------------+---------------------------------------------------------------
           . |      6,249  -4681.199  -4386.222       8   8788.443   8842.365
-----------------------------------------------------------------------------
Note: BIC uses N = number of observations. See [R] IC note.

.  logit viq220 c.ridageyr i.ridreth1 i.riagendr c.indfmpir, or

Iteration 0:  Log likelihood = -4260.6598  
Iteration 1:  Log likelihood = -3949.4647  
Iteration 2:  Log likelihood = -3948.0389  
Iteration 3:  Log likelihood = -3948.0387  

Logistic regression                                     Number of obs =  6,249
                                                        LR chi2(7)    = 625.24
                                                        Prob > chi2   = 0.0000
Log likelihood = -3948.0387                             Pseudo R2     = 0.0734

------------------------------------------------------------------------------
      viq220 | Odds ratio   Std. err.      z    P>|z|     [95% conf. interval]
-------------+----------------------------------------------------------------
    ridageyr |   .9780431   .0012665   -17.15   0.000      .975564    .9805284
             |
    ridreth1 |
          2  |   .8891554   .1496069    -0.70   0.485     .6393767    1.236513
          3  |   .6051742   .0454583    -6.69   0.000     .5223259    .7011634
          4  |   .8115435   .0642782    -2.64   0.008     .6948522    .9478315
          5  |   .5860531   .0821287    -3.81   0.000     .4452979    .7712998
             |
  2.riagendr |   .5974353   .0324396    -9.49   0.000     .5371212    .6645221
    indfmpir |   .8927698   .0158011    -6.41   0.000     .8623312    .9242828
       _cons |   7.516118   .6598894    22.97   0.000     6.327921    8.927424
------------------------------------------------------------------------------
Note: _cons estimates baseline odds.

. 
end of do-file

. 

```

```stata

 estout m1 m2 m3, stats(aic N r2_p) eform
 
```

to produce

```stata

.  estout m1 m2 m3, stats(aic N r2_p) eform

---------------------------------------------------
                       m1           m2           m3
                        b            b            b
---------------------------------------------------
viq220                                             
ridageyr          .975618     .9776661     .9780431
1.ridreth1                           1            1
2.ridreth1                     .854417     .8891554
3.ridreth1                    .5122235     .6051742
4.ridreth1                    .7688672     .8115435
5.ridreth1                    .5209638     .5860531
1.riagendr                           1            1
2.riagendr                    .6058933     .5974353
indfmpir                                   .8927698
_cons            3.532101     6.281091     7.516118
---------------------------------------------------
aic              8477.679     8289.925     7912.077
N                    6547         6547         6249
r2_p             .0497642     .0719403     .0733739
---------------------------------------------------

. 
end of do-file

. 

```

## citation:

"https://www.stata.com/support/faqs/statistics/outcome-does-not-vary/"
"http://repec.org/bocode/e/estout/hlp_estout.html#cells"
"https://stats.oarc.ucla.edu/stata/faq/how-can-i-use-estout-to-make-regression-tables-that-look-like-those-in-journal-articles/"


#### Problem 1(d)

Based on the result from last part, we can see a p-value of 0.00 in the chart. So we reject the null hypothesis since p<0.05. Thus, the odds of men and women being wears of glasses/contact lenses for distance vision differs. Based on the Logistic regression model, we see that the p-value for gender is 0 which is less than 0.05 so we reject the null hypothesis for it and therefor the proportion of wearers of glasses/contact lenses for distance vision differs between men and women. 


#### Problem 2(a)
```{r}
library(DBI)
library(RSQLite)

sakila <- dbConnect(RSQLite::SQLite(), "~/Desktop/sakila_master.db")
dbListTables(sakila)
```


```{r}

dbGetQuery(sakila, "
Select Count(*), language_id
from film
GROUP BY language_id
")

##used original language_id to make sure it is correct
dbGetQuery(sakila, "
Select Count(*), original_language_id
from film
GROUP BY original_language_id
")
```

Based on the result, it seems like English is the most common used language for the films. (I double check with the original one and it seems like English is the only one)



#### Problem 2(b)
## way 1 (SQL)
```{r}
dbGetQuery(sakila, "
SELECT c.category_id AS CID, c.name AS NAME, COUNT(f.category_id) AS num_cat
  FROM category AS c
       INNER JOIN film_category AS f ON c.category_id = f.category_id
 GROUP BY c.category_id
 ORDER BY -num_cat
")
```

Based on the result, it looks like sports movie is the most common in the data and there are 74 movies about sports.


## way 2 (R)
```{r}
df1_names <- dbGetQuery(sakila, "SELECT category_id, name FROM category")
df2 <- dbGetQuery(sakila, "SELECT category_id, film_id FROM film_category")
genre_num <- tail(names(sort(table(df2$category_id))), 1)
genre_num
num_of_ocu <- length(which(df2$category_id == 15))
num_of_ocu
```


Based on the result, it looks like sports movie is the most common in the data and there are 74 movies about sports.


#### Problem 2(c)
## way 1 (SQL)
```{r}
dbGetQuery(sakila, "
SELECT a.country_id AS COUNTRYID, a.country AS NAME, COUNT(b.country) AS num_count
  FROM country AS a
       INNER JOIN customer_list AS b ON a.country = b.country
 GROUP BY a.country
 HAVING num_count == 9
 ORDER BY -num_count
")
```

Based on the result, only the United Kingdom has 9 customers.


## way 2 (R)
```{r}
df4 <- dbGetQuery(sakila, "SELECT country FROM customer_list")
which(table(df4$country)==9)
```

Based on the result, only the United Kingdom has 9 customers.


#### Problem 3(a)

```{r}
us_500 <- read.csv("~/Desktop/us-500.csv")
num_tld <- length(grep(".net", us_500$email))
num_emails <- length(us_500$email)
proportion <- num_tld/num_emails
proportion
```

The proportion of email addresses are hosted at a domain with TLD “.net" is 14.6% based on the result.



#### Problem 3(b)
```{r}
num_na <- length(grep("[^a-zA-Z\\d\\s:\\@\\.]+", us_500$email))
proportion2 <- num_na/num_emails
proportion2
```

Based on the result, 24.8% of email addresses have at least one non alphanumeric character in
them.


#### Problem 3(c)
```{r}
phones_col <- rbind(us_500$phone1,us_500$phone2)
phones_count <- substr(phones_col, 0, 3)
table <- table(phones_count)
names(table)[which.max(table)]
```

The most used area code is 973.




#### Problem 3(d)
```{r}
df_a <- data.frame(gsub(".*\\D", "", us_500$address))
## The following step is to remove all the empty "" from the data frame above
clean_data <- df_a[!(is.na(df_a) | df_a==""), ]
## Taking the log value of cleaned data and make them numeric
nume_data <- log(as.data.frame(sapply(clean_data, as.numeric)))
hist(nume_data$`sapply(clean_data, as.numeric)`, main  = "Histogram", xlab = "log(Apartment Number)")
```



#### Problem 3(e)
```{r}
first_let <- substr(clean_data, 0, 1)
table(first_let)
max(names(table(first_let)))
```


Based on the result, it seems like 9 appears as the leading significant digit and then 1. From 2-8, it doesn't look like it's in a decreasing order. Therefore, the apartment numbers does not appear to follow Benford’s law and the apartment numbers would not pass as real data.



#### Problem 3(f)
```{r}
street_num <- sub("\\D+.*", "", us_500$address)
last_digit <- substr(street_num, nchar(street_num), nchar(street_num))
table(last_digit)
max(names(table(last_digit)))
```


Based on the result, it seems like 9 is again the most significant digit and on the other side 1 is the lowest(we are not counting 0 since we are looking at the first significant digit). Therefore, the last digit of street numbers does not appear to follow Benford’s law and the street numbers would not pass as real data.





