---
title: "STATS 506 HW6"
format: html
editor: visual
embed-resources: true
code-fold: true
---

Github Link: 
#### Problem 1

```{r}
## Without any parallel process
library(boot)
load("~/Desktop/nycflights13_4/data/flights.rda")
flights <- na.omit(flights)
flights$air_time <- as.numeric(flights$air_time)
flights$origin <- as.factor(flights$origin)
flights$dest <- as.factor(flights$dest)
```


```{r}
#' A function called boot_fn to use as statistic in the boot later to return the mean air_time
#'
#' @param data 
#' @param indices 
#'
#' @return mean air_time

boot_statistics <- function(data, indices) {
  subset_data <- data[indices,]
  mean_air_time <- tapply(subset_data$air_time, subset_data$origin, mean)
  return(mean_air_time)
}

set.seed(123)

resamples <- boot(flights, boot_statistics, strata = flights$dest, R = 1000)
resamples

## calculate my 95% CI using standard error
se <- apply(resamples$t, 2, sd)
CI_lwb <- resamples$t0 - 1.96 * se
CI_upb <- resamples$t0 + 1.96 * se
mean_estimates <- resamples$t0

final_table <- data.frame(
  Mean_estimates = mean_estimates,
  Standard_error = se,
  CI_lower = CI_lwb,
  CI_upper = CI_upb
)

print(final_table)
```




```{r}
## With parallel processing
library(future)
plan(multisession)
options(future.rng.onMisuse = "ignore")


## same function as above
boot_statistics2 <- function(data, indices) {
  subset_data <- data[indices,]
  mean_air_time <- tapply(subset_data$air_time, subset_data$origin, mean)
  return(mean_air_time)
}

## bootstrap using future package
set.seed(123)

resamples_p <- vector("list", 1000)

for (i in 1:1000) {
  p <- future({ 
    boot_statistics2(flights, sample(nrow(flights), replace = TRUE)) 
  })
  resamples_p[[i]] <- p
}


resamples_p_df <- data.frame()
for(i in 1:length(resamples_p)){
  df <- as.data.frame(t(value(resamples_p[[i]])))
  resamples_p_df <- rbind(resamples_p_df, df)
}

## calculate 95%CI using standard error
se_p <- apply(resamples_p_df, 2, sd)
mean_estimates_p <- colMeans(resamples_p_df)
CI_lwb_p<- mean_estimates_p - 1.96 * se_p
CI_upb_p <- mean_estimates_p + 1.96 * se_p

final_table_p <- data.frame(
  Mean_estimate = mean_estimates_p,
  Std_error = se_p,
  CI_lower = CI_lwb_p,
  CI_upper = CI_upb_p
)

print(final_table_p)

```


Based on two tables, I think the mean estimate is around the same. The only differences is in the decimals and I don't think it matters too much. But for standard errors, it looks the the one using future package has a higher number comparing to the one using non parallel processing. This leads to and higher CI lower bond for the one without any parallel processing and a higher CI upper bond for the one using parallel processing.





　

